### 2023-07-27 draft.tex
First step for a draft

\documentclass{article}

\usepackage[utf8x]{inputenc}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{xcolor}

\usepackage{listings}

\input{DIALEKTO}

\title{Sequent calculus style in Dedukti}
\author{Simon Guilloud and Amélie Ledein}
\date{}

\begin{document}

	\maketitle

	\begin{abstract}
		This draft presents a sequent calculus in Dedukti in order to, in the future, translate and recheck LISA proofs.
	\end{abstract}

	Computer science is spreading increasingly in everyday human life: medecine, transport, communication or also teaching. In some circumstances, software is the critical part of a process as human lives are at stake.
	For that reason, research at the intersection between mathematics, computer science and logic develops various formal tools
	% Formal tools exist and are based on different logics.
	based on different logics to increase the confidence over softwares.
	Such tools can be SAT or SMT solvers, 
	%proof assistants as 
	Lean, Coq, Agda, B, HOL Light, Isabelle, PVS, Metamath, K, etc.
	To deeply study and understand the logical differences between all these formal tools, some logical frameworks have been elaborated such as Dedukti.

	Dedukti is a logical framework based on the $\lambda\Pi$-calculus modulo theory, a typed $\lambda$-calculus with dependent types and rewriting rules.
	The flexibility and the expressivity of this logical framework suggest that Dedukti is a good candidate to allow interoperability of proofs. \\

	%\begin{itemize}
	%	\item justification de Dedukti
	%	\item contribution
	%\end{itemize}

	After presenting Dedukti as well as sequent calculus, this article proposes an encoding of sequent calculus

\newpage

	\section{Preliminaries}

	\subsection{The spirit of Dedukti encodings}

	using dependent types, Curry-De Bruijn-Howard isomorphism and rewriting rules.

	\paragraph{First-order logic.} First, we In order to define FOL in Dedukti.

	The idea is the symbols \textsf{Term} and \textsf{Formula} are types as their type is TYPE. Moreover, these symbols are declared as constant to avoid rewriting.

	First, we define the type of terms
	This type can be extended

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={8-8}]{SimpleVersion/FOL.lp}

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={4-4}]{SimpleVersion/FOL.lp}

	A formula respects the following grammar where $t ::= x~|~...~~$: \\
	$\phi ::= \textsf{Top}~|~\textsf{Bot}~|~\phi~\land~\phi~|~\phi~\lor~\phi~|~\phi~\textsf{Implies}~\phi~|~\textsf{Forall}~f~|~\textsf{Exists}~f~|~\textsf{Neg}~\phi~|~\phi~\textsf{Iff}~\phi~|~\textsf{ExistsUniq}~f~|~t~\textsf{Equality}~t$

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={13-14,16-18,20-21,23-25}]{SimpleVersion/FOL.lp}

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={11-11}]{SimpleVersion/FOL.lp}

	The combinaison of these symbols defines formulas.
	In order to do proofs, .

	\begin{lstlisting}[language=Dialekto, firstnumber=last]
symbol Prf : Formula → TYPE;
	\end{lstlisting}

	\begin{lstlisting}[language=Dialekto, firstnumber=last]
symbol id-ax A : Prf (A Implies A);
symbol iff-ax A B : Prf (A Implies B) → Prf (B Implies A) → Prf (A Iff B);
	\end{lstlisting}

	\begin{lstlisting}[language=Dialekto, firstnumber=last]
symbol equiv-prf A : Prf (A Iff A) := iff-ax A A (id-ax A) (id-ax A);
	\end{lstlisting}

\newpage

	%\paragraph{Programming language.}
	\paragraph{Computation side.}

	Now we can define natural numbers.

	% Natural numbers
	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={20-20, 22-23}]{SimpleVersion/prelude.lp}

	This declaration is very similar to the definition of an inductive type.

	Thanks to rewriting rules, it is possible to leave the rewriting engine of Dedukti to do computations.

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={1-1}]{SimpleVersion/prelude.lp}

	Interpretation of set codes in TYPE

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={5-5}]{SimpleVersion/prelude.lp}

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={29-29,31-31}]{SimpleVersion/prelude.lp}

	%List
	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={7-9,11-11}]{SimpleVersion/prelude.lp}

	% Append
	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={13-13,15-17}]{SimpleVersion/prelude.lp}

	% Map and reverse
	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={33-39}]{SimpleVersion/prelude.lp}

	List

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={104-108}]{SimpleVersion/prelude.lp}

	\paragraph{To go further.}
	We have just presented the basic concepts in order to encode a logic into Dedukti.
	Various logics have been already defined

	Calculus of Construction~\ref{},

	predicate subtyping~\ref{}, universe polymorphism~\ref{}

	However logics based on sequent calculus are less explored.

	\subsection{Sequent calculus}

\newpage

	\section{Encoding of Sequent calculus in Dedukti}

	We have already defined first-order formula.

	As we have done for natural numbers, we declare a type of formula at the level of Set in order to define a list of formulas

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={5-6}]{SimpleVersion/FOL.lp}

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={28-29}]{SimpleVersion/FOL.lp}

	\lstinputlisting[language=Dialekto, firstnumber=last, linerange={31-31}]{SimpleVersion/FOL.lp}

	\section{Link with other logics}

	\subsection{Minimalistic}

	\subsection{Intuitionistic}

	\subsection{Classic}

	\section{Alternative}

	A sequent is a pair of finite lists which can be noted as follows: $H_1,...,H_m \vdash C_1,...,C_n$, where $H_i$ are hypotheses and $C_j$ are conclusions, $m \ge 0$ and $n \ge 0$.

	The usual interpretation of a sequent is
	$H_1 \land ... \land H_m \Rightarrow C_1 \lor ... \lor C_j$.

	As the logical connectives $\land$ and $\lor$ are commutative, it is also suitable to consider that a sequent is a pair of sets.

	However, as a set respects the properties of associative, commutative, unity and idempotence, it is less convenient to manage it with rewriting rules.
	To do that, we need to increase the expressivity of considered rewriting systems reasoning modulo ACUI.

	We implement

	In practice,

	AC ?

	\section{LISA}

	LISA is ...

	The proposed encoding can be seen as the first step to translate LISA into Dedukti as the encoded sequent calculus is similar to the one used for LISA.
	For now, LISA does not generate any kind of hints that are necessary to translate and recheck LISA proofs.

	The next step of this work is to instrument LISA in order to develop a translator from LISA to Dedukti.

	\section{Conclusion}

\end{document}